---
title: "appropriate_abx_public"
author: "Julie Yarahuan"
date: '2022-09-08'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Define Variables

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. 

```{r,echo=FALSE, message=FALSE}
#Load Libraries
library(tidyr)
library(tibble)
library(dplyr)
library(readr)
library(lubridate)
library(stringr)
library(data.table)
library(tinytex)
```

```{r, echo=FALSE, message=FALSE}
##Input your variables here##

#File path variable: Input your file path here
abxfile="/Volumes/dmcommon/_DM.Private/CI Program Faculty/+Fellow Franchise/Yarahuan/Protected/Antibiotics.csv"
culturefile="/Volumes/dmcommon/_DM.Private/CI Program Faculty/+Fellow Franchise/Yarahuan/Antibiogram/Positive Blood or Respiratory Cultures with organism.csv"

#Load Datasets
abxdata <- read_csv(abxfile)
culturedata<-read_csv(culturefile)

#column names for antibiotics table, make sure your data is in the below order: 
#1=contact serial number (encounter ID), 2=MRN, 3=date of birth, 4= Admit DateTime, 5=Service Abbreviation, 6=Service description, 7=ICD10 code, 8=ICD 10 description,
#9=Medication Order DateTime, 10=Medication route,11=Medication in Orderset/Powerplan, 12=allergy substance, 13=allergy type

#rename columns in abx data
abxdata<-abxdata %>%
  rename(CSN=1) %>%
  rename(MRN=2) %>%
  rename(DOB=3) %>%
  rename(AdmitDT=4) %>%
  rename(Service=5) %>%
  rename(ServiceDesc=6) %>%
  rename(ICD=7) %>%
  rename(ICDdesc=8) %>%
  rename(MedDT=9) %>%
  rename(Med=10) %>%
  rename(MedRoute=11) %>%
  rename(Orderset=12) %>%
  rename(Allergy=13) %>%
  rename(AllergyType=14)

#column names for culture table, make sure your data is in the following order
#1=MRN, 2=lab order, 3=Collection DT, 4= Organism, 5=suceptibility (eg. susceptble or resistant), 6= antibiotic susceptbility (antibiotic name), 7=susceptibility result DT

#rename columns in culture data
culturedata<-culturedata %>%
  rename(MRN=1) %>%
  rename(LabOrder=2)%>%
  rename(CollectionDT=3)%>%
  rename(Organism=4)%>%
  rename(Susceptibility=5)%>%
  rename(SusceptibilityAntibiotic=6)%>%
  rename(SusceptibilityDT=7)

```

```{r,echo=FALSE, message=FALSE}
#Join Datasets

#reformat MRN as numeric in culture dataset
culturedata$MRN <- as.numeric(culturedata$MRN)

#join abx with cultures
abxculturedata= abxdata %>% left_join(culturedata,by="MRN")
```

```{r,echo=FALSE, message=FALSE}
#Show table to ensure columns & data are loaded and joined correctly
head(abxculturedata)
```

```{r, echo=FALSE}
################################# Clean Dataset ########################################

#reformat date in culture dataset
abxculturedata$CollectionDT<-mdy_hms(abxculturedata$CollectionDT)
abxculturedata$SusceptibilityDT<-mdy_hms(abxculturedata$SusceptibilityDT)

#reformat dates for admit DT and DOB
abxculturedata$AdmitDT<-mdy_hms(abxculturedata$AdmitDT, truncated = 1)
abxculturedata$DOB<-mdy(abxculturedata$DOB)

#Calculate age in years at time of admission
abxculturedata <- abxculturedata %>%
  mutate(age=year(AdmitDT)-year(DOB))

#calculate age in months at time of admission
abxculturedata<-abxculturedata %>%
  mutate(agemonths=interval(DOB,AdmitDT) %/% months(1))

#define inpatient vs ED encounters
abxculturedata<-abxculturedata %>%
  mutate(servicelocation= if_else(Service=="EMR","Emergency","Inpatient",missing=NULL))

#make text columns lowercase
abxculturedata$ICDdesc=tolower(abxculturedata$ICDdesc)
abxculturedata$Med=tolower(abxculturedata$Med)
abxculturedata$Allergy=tolower(abxculturedata$Allergy)
abxculturedata$SusceptibilityAntibiotic=tolower(abxculturedata$SusceptibilityAntibiotic)
abxculturedata$Susceptibility=tolower(abxculturedata$Susceptibility)

#Count unique starting CSNs
csnunique <- unique(abxculturedata$CSN)

paste("Total starting encounters: ",length(csnunique))

```

## Define  guideline exclusions and guideline choices 
```{r, echo=FALSE}
#Select services of interest
servicesofinterest<-c('PMA','PMB','PMC','PMD','CCS','PUL','EMR','ADL') 

#Ages to exclude (e.g. this guideline excludes children less than 3 months old)
minagemonths=3

#Co-morbid conditions to exclude (make all lowercase)
exclusiondiagnoses=c("cystic fibrosis","sickle-cell","hb-ss","sickle cell","leukemia","lymphoma",'neoplasm','transplant',"tuberculosis",
                     "tracheostomy","end stage renal","immunodeficiency",'immunodeficiencies','hypogammaglobulinemia','di george',
                     "rheumatoid arthritis","lupus","muscular dystrophy","spinal muscular atrophy","hypoplastic left heart syndrome",
                     "pulmonary hypertension","cardiomyopathy","heart failure")


#filter out some common infections based on ICD prefixes
#AO: exclude infectious colitis, "B0*":exclude many viral infections, "J95.851":exclude ventilator assc PNA, "M00*":septic arthritis,  "M01*": septic arthritis,"N39*": exclude UTI
exclusioninfections<-c("A0*","B0*","J95.851", "M00*","M01*","N39*")
exclusioninfections2=c("abscess","cellulitis", "osteomyelitis",
                     "mastoiditis","lymphangitis","sinusitis","streptococcal pharyngitis","impetigo","lymphadenitis",
                     "omphalitis","aspiration","arthritis due to other bacteria","myositis","sinusitis",'pyelonephritis')

#filter out certain antimicrobials
exclusionantimicrobials<- c('oseltamivir','bacitracin','nystatin','valacyclovir','hydroxychloroquine','erythromycin')

#identify list of pneumonia ICD codes of interest (these are the codes of pneumonia that you are INCLUDING in your table)
pnalist<-c('J13', 'J13', 'J14', 'J14', 'J15.0', 'J15.0', 'J15.1', 'J15.1', 'J15.20', 'J15.20', 'J15.211', 'J15.211', 'J15.212', 'J15.212', 'J15.29', 'J15.29', 'J15.3', 'J15.3', 'J15.4', 'J15.4', 'J15.5', 'J15.5', 'J15.6', 'J15.6', 'J15.7', 'J15.7', 'J15.8', 'J15.8', 'J15.9', 'J15.9', 'J16.0', 'J16.0', 'J16.8', 'J17', 'J18.0', 'J18.1', 'J18.8', 'J18.9', 'J85.1', 'J95.851', 'J95.851')

#define first line choices 
#####first line is ampicillin or amoxicillin
#####if PCN allergy then, ceftriaxone, clindamycin, cefdinir, levofloxacin
#####if cephalosporin allergy then, levofloxacin, clindamycin
#first line atypical coverage is azithromycin
######if azithromycin allergy then levofloxacin or doxycycline
firstline=c("amoxicillin","ampicillin")
PCNallergy=c("amoxicillin","ampicillin",'penicillin','augmentin','penicillins','amoxicillin-clavulanate') #list of meds that count as PCN allergy
ifPCNallergy=c("ceftriaxone","clindamycin","cefdinir",'levofloxacin')
cephalosporinallergy=c("cefdinir","cefotaxime","ceftriaxone","cephalexin","cephalosporins","cefuroxime","cefazolin","cefadroxil","cefotetan","cefoxitin","ceftazidime","cefpodoxime","cefixime","cefepime","ceftaroline") #list of meds that count as cephalosporin allergy
ifcephalosporinallergy=c("levofloxacin","clindamycin")
atypicalfirstline="azithromycin"
azithroallergy="azithromycin"
atypicalwithazithroallergy=c("levofloxacin","doxycycline")

#if over 4 then azithromycin is potentially appropriate
ageforazithro=4 

```

```{r, echo=FALSE}
################################# Define culture exclusions ########################################

#pcnresistance list
narrowpcnsusceptibilitylist<-c("ampicillin","oxacillin",'Penicillin (meningitis)','Penicillin (non-meningitis)')
broadpcnsuscetiblitylist<-c("ampicillin","oxacillin",'ampicillin/sulbactam','Penicillin (meningitis)','Penicillin (non-meningitis)')
```

# Output


## Summary of reasons for exlusion
```{r, echo=FALSE}
################################# Apply guideline exclusions ########################################
#eliminate patients less than your min exclusion age
abxculturedata <- abxculturedata %>%
  mutate(reasonforexclusions= if_else(
    agemonths>=minagemonths,"included","below min age"))

#select services of interest
abxculturedata <- abxculturedata %>%
  mutate(reasonforexclusions= if_else(
    (Service %in% servicesofinterest) &reasonforexclusions=="included","included",
    if_else(reasonforexclusions!="included",reasonforexclusions,"excluded service")))

#filter diagnoses that have any of the excluded diagnoses
excludelist<-abxculturedata %>%
  select(CSN) %>%
  filter(str_detect(abxculturedata$ICDdesc, paste(exclusiondiagnoses,collapse='|')))

abxculturedata <- abxculturedata %>%
  mutate(reasonforexclusions= if_else(
    CSN %in% excludelist$CSN & reasonforexclusions=="included","excluded diagnosis",
    if_else(reasonforexclusions!="included",reasonforexclusions,"included")))

#filter diagnoses that have any other common infections
excludeinfectionlist<-abxculturedata %>%
  select(CSN) %>%
  filter(abxculturedata$ICD %like% paste(exclusioninfections,collapse = "|"))

#apply exclude list
abxculturedata <- abxculturedata %>%
  mutate(reasonforexclusions= if_else(
    CSN %in% excludeinfectionlist$CSN & reasonforexclusions=="included","excluded other common infection",
    if_else(reasonforexclusions!="included",reasonforexclusions,"included")))

#common infections exclusions list 2
excludeinfectionlist2<-abxculturedata %>%
  select(CSN) %>%
  filter(str_detect(abxculturedata$ICDdesc, paste(exclusioninfections2,collapse='|')))

abxculturedata <- abxculturedata %>%
  mutate(reasonforexclusions= if_else(
    CSN %in% excludeinfectionlist2$CSN & reasonforexclusions=="included","excluded other common infection",
  if_else(reasonforexclusions!="included",reasonforexclusions,"included")))


#exclude selected antimicrobials from med list
excludemedlist<-abxculturedata %>%
  select(CSN) %>%
  filter(abxculturedata$Med %like% paste(exclusionantimicrobials,collapse="|"))

abxculturedata <- abxculturedata %>%
  mutate(reasonforexclusions= if_else(
    CSN %in% excludemedlist$CSN & reasonforexclusions=="included","did not receive antibiotic",
  if_else(reasonforexclusions!="included",reasonforexclusions,"included")))

#pna subset
pnasub<-abxculturedata %>%
  select(CSN) %>%
  filter(abxculturedata$ICD %in% pnalist)

abxculturedata <- abxculturedata %>% 
  mutate(reasonforexclusions= if_else(CSN %in% pnasub$CSN & reasonforexclusions=="included","included",
                                      if_else(reasonforexclusions!="included",reasonforexclusions,"no common pnuemonia ICD code")))


exclusionsummarytable<-setNames(aggregate(CSN ~ reasonforexclusions, abxculturedata, function(x) length(unique(x))),c("Reason For Exclusion","Count Excluded Encounters"))


grandtotal<-c("Grand Total",length(unique(abxculturedata$CSN)))
exclusionsummarytable<-rbind(exclusionsummarytable,grandtotal)


exclusionsummarytable

```
```{r, echo=FALSE,message=FALSE, results=FALSE}
################################# Apply guidelines to assess antibiotic choice ########################################

#find first abx ordered at encounter
abxchoice <- abxculturedata %>%
  group_by(CSN) %>%
  filter(
    MedDT== min(MedDT, na.rm = T) & reasonforexclusions=="included") 

#create a blank column to store reasons abx are classified as incorrrect: 
abxchoice <- abxchoice %>%
  add_column(label = NA)

#correct based on defined guidelines in definitions section
abxchoice<-abxchoice %>%
  ungroup()%>%
  mutate(correctinitialabx=case_when(
    Med %in% firstline ~1,
    (Allergy %in% PCNallergy & Med %in% ifPCNallergy)~1,#if allergy to pencillin, then Ceftriaxone, clinda, cefdinir or levofloxacin are correct
    (Allergy %in% cephalosporinallergy & Med %in% ifcephalosporinallergy)~1,#if cephalosporin allergy then clinda or levofloxacin
    (age>ageforazithro & Med==atypicalfirstline)~1,#if over 4 then azithromycin correct,
    (age>ageforazithro & Allergy==azithroallergy & Med %in% atypicalwithazithroallergy)~1, #if over 4 and has azithro allergy
    TRUE~0
  ))

#add reasons for being incorrect
abxchoice<-abxchoice %>%
  ungroup()%>%
  mutate(label=case_when(
    Med %in% firstline ~"First Line Choice",
    (Allergy %in% PCNallergy & Med %in% ifPCNallergy)~"Correct substitute i/s/o PCN allergy",
    (Allergy %in% cephalosporinallergy & Med %in% ifcephalosporinallergy)~"Correct substitute i/s/o cephalosporin allergy",
    (age>ageforazithro & Med==atypicalfirstline)~"Appropriate age for Azithro use",
    (age>ageforazithro & Allergy==azithroallergy & Med %in% atypicalwithazithroallergy)~"Appropriate age for azithro with appropriate azithro subs i/s/o   Azithro allergy",
    TRUE~"Likely incorrect abx"
  ))

#take max of choice per patietn (this eliminates rows that look false [e.g. patient with multiple drug allergies, recorded as 'incorrect' for one allergy but correct for another])
maxabxchoice<-abxchoice %>% group_by(CSN) %>% top_n(1,correctinitialabx)
  
#summary of correct vs incorrect choices
summarytable1<-maxabxchoice %>%
  group_by(correctinitialabx)%>%
  summarise(count=n_distinct(CSN))%>%
  ungroup()

correctlabels<-c("Incorrect", "Correct")

cbind(correctlabels,summarytable1)

#Summary of labels (please note that the "correct" count may be higher on this because many patients are ordered for both azithromycin and a first line antibiotic at the same time so the same patient will be in two separate label categories. This does not change the correct/incorrect count)
maxabxchoice%>%
  group_by(label)%>%
  summarise(count=n_distinct(CSN))%>%
  ungroup()

```

## Summary of Correct/Incorrect Orders with Labels [Pre-NLP Analysis]
```{r,echo=FALSE}
###################### Culture Results Analysis ##################

#find history of positive culture preceeding abx order date
maxabxchoice <- maxabxchoice %>%
  mutate(hxposculture=NA)

maxabxchoice<-maxabxchoice %>%
  mutate(hxposculture= if_else(CollectionDT<AdmitDT,1,0,missing=0)) #1 if culture collected prior to admission

maxabxchoice<-maxabxchoice %>%
  mutate(correctinitialabx=case_when(
    (correctinitialabx==0 & hxposculture==1 & Susceptibility=='Resistant'& SusceptibilityAntibiotic %in% broadpcnsuscetiblitylist   & (Med=='ceftriaxone'|Med=='clindamycin'|Med=='cefdinir'|Med=='cephalexin')) ~1,
     (correctinitialabx==0 &hxposculture==1 & Susceptibility=='Resistant'& SusceptibilityAntibiotic %in% narrowpcnsusceptibilitylist   & (Med=='ceftriaxone'|Med=='cephalexin'|Med=='cefdinir'|Med=='amoxicillin-clavulanate'|Med=='ampicillin-sulbactam')) ~1,
    TRUE~correctinitialabx
  ))

maxabxchoice<-maxabxchoice %>%
  mutate(label=case_when(
    (correctinitialabx==0 & hxposculture==1 & Susceptibility=='Resistant'& SusceptibilityAntibiotic %in% broadpcnsuscetiblitylist   & (Med=='ceftriaxone'|Med=='clindamycin'|Med=='cefdinir'|Med=='cephalexin')) ~"Prior Resistant Culture",
     (correctinitialabx==0 &hxposculture==1 & Susceptibility=='Resistant'& SusceptibilityAntibiotic %in% narrowpcnsusceptibilitylist   & (Med=='ceftriaxone'|Med=='cephalexin'|Med=='cefdinir'|Med=='amoxicillin-clavulanate'|Med=='ampicillin-sulbactam')) ~"Prior Resistant Culture",
    TRUE~label
  ))

#take max of choice per patietn
maxabxchoice<-abxchoice %>% group_by(CSN) %>% top_n(1,correctinitialabx)

#summary of correct vs incorrect choices
summarytable2<-maxabxchoice %>%
  group_by(correctinitialabx)%>%
  summarise(count=n_distinct(CSN))%>%
  ungroup()

correctlabels<-c("Incorrect", "Correct")

cbind(correctlabels,summarytable2)

aggregate(CSN ~ label, maxabxchoice, function(x) length(unique(x))) 
```



```{r, echo=FALSE}
################################# get list of all remaining patients who are incorrect for NLP below ########################################

#get CSN list of remaining "incorrect" patients
csnlist<-maxabxchoice[maxabxchoice$correctinitialabx == 0, ]
csnlist<-unique(csnlist$CSN)

#save CSN list as CSV

write.csv(csnlist,"/Volumes/dmcommon/_DM.Private/CI Program Faculty/+Fellow Franchise/Yarahuan/Protected/CSNList.csv", row.names = FALSE)
```

```{r,echo=FALSE, message=FALSE}
################################# Start NLP ########################################################

#File path variable: Input your file path to note contents file here
notefile="/Volumes/dmcommon/_DM.Private/CI Program Faculty/+Fellow Franchise/Yarahuan/Protected/NoteContents.csv"

#Load Datasets
notedata <- read_csv(notefile)

#NoteData Columns
#1=CSN, 2=NoteDT, 3=NoteType, 4=NoteText

#rename columns
notedata<-notedata %>%
  rename(CSN=1) %>%
  rename(NoteDT=2) %>%
  rename(NoteType=3) %>%
  rename(NoteText=4)

#make text columns lowercase
notedata$NoteType=tolower(notedata$NoteType)
notedata$NoteText=str_replace_all(notedata$NoteText, "[^[:alnum:]]", " ")
notedata$NoteText=tolower(notedata$NoteText)


```


```{r, echo=FALSE,message=FALSE}
################################# Create your NLP phrases ########################################################

treatmentfailurephrases=c("treatment failure","failed outpatient treatment","failed outpatient therapy","failed outpatient antibiotics",
                          "failure of outpatient treatment", "failure of outpatient therapy", "failure of outpatient amoxcillin"
                          ,"failure of outpatient HD amoxicillin","failure of outpatient high dose amoxicillin","failure of outpatient high-dose amoxicillin"
                          ,"despite amoxicillin","despite outpatient antibiotics", "despite outpatient therapy", "despite outpatient treatment",
                          "in spite of outpatient treatment","s p outpatient therapy","s p outpatient treatment","s p outpatient antibiotics",
                          "s p outpatient amoxicillin","sp outpatient therapy","sp outpatient treatment","sp outpatient antibiotics",
                          "sp outpatient amoxicillin", "s/p outpatient therapy","s/p outpatient treatment","s/p outpatient antibiotics",
                          "s/p outpatient amoxicillin","started on amoxicillin as outpatient", "started on HD amoxicillin as outpatient",
                          "started on high dose amoxicillin as outpatient","started on amoxicillin as an outpatient", 
                          "started on HD amoxicillin as an  outpatient","started on high dose amoxicillin as an outpatient")

aspirationphrases=c("aspiration pneumonia","likely aspiration", "aspiration event")

unimmunizedphrases=c("unimmunized","unvaccinated","partially vaccinated","partially immunized")

```

## Summary of NLP findings
```{r,echo=FALSE,message=FALSE}
################################# Apply NLP phrases ########################################################

#apply treatment failure phrases
treatmentfailure<-notedata %>%
  select(CSN) %>%
  filter(str_detect(notedata$NoteText, paste(treatmentfailurephrases,collapse='|')))

#apply aspiration phrases
aspiration<-notedata %>%
  select(CSN) %>%
  filter(str_detect(notedata$NoteText, paste(aspirationphrases,collapse='|')))

#apply unimmunized phrases
unimmunized<-notedata %>%
  select(CSN) %>%
  filter(str_detect(notedata$NoteText, paste(unimmunizedphrases,collapse='|')))

#get all CSNs from treatment failure list
treatmentfailurecsnlist<-unique(treatmentfailure$CSN)
aspirationcsnlist<-unique(aspiration$CSN)
unimmunizedcsnlist<-unique(unimmunized$CSN)


#if CSN in treatmnet failure list, assign to value of 2
maxabxchoice<-maxabxchoice %>%
  ungroup()%>%
  mutate(correctinitialabx= if_else(maxabxchoice$CSN %in% treatmentfailurecsnlist
                                    ,2,correctinitialabx,correctinitialabx))

maxabxchoice<-maxabxchoice %>%
  ungroup()%>%
  mutate(label= if_else(maxabxchoice$CSN %in% treatmentfailurecsnlist
                                    ,"outpatient treatment failure",label,label))

#if CSN in aspiration list, assign to value of 3
maxabxchoice<-maxabxchoice %>%
  ungroup()%>%
  mutate(correctinitialabx= if_else(maxabxchoice$CSN %in% aspirationcsnlist
                                    ,3,correctinitialabx,correctinitialabx))

maxabxchoice<-maxabxchoice %>%
  ungroup()%>%
  mutate(label= if_else(maxabxchoice$CSN %in% aspirationcsnlist
                                    ,"aspiration pneumonia",label,label))

#if CSN in unimmunized list, assign to value of 4
maxabxchoice<-maxabxchoice %>%
  ungroup()%>%
  mutate(correctinitialabx= if_else(maxabxchoice$CSN %in% unimmunizedcsnlist
                                    ,4,correctinitialabx,correctinitialabx))
maxabxchoice<-maxabxchoice %>%
  ungroup()%>%
  mutate(label= if_else(maxabxchoice$CSN %in% unimmunizedcsnlist
                                    ,"unimmunized",label,label))

#take max of choice per patietn
maxabxchoice<-maxabxchoice %>% group_by(CSN) %>% 
  mutate(finalcorrectinitialabx= max(correctinitialabx)) %>%
  ungroup()

maxabxchoice %>%
  group_by(correctinitialabx)%>%
  summarise(count=n_distinct(CSN))%>%
  ungroup()

```
## Summary of NLP after applying additional guidelines for unimunized patients
```{r,echo=FALSE,message=FALSE}
################################# Apply additional guidelines based on above NLP ########################################################

#define guidelinechoices (if hospitalized & unimmunized or failed outpatient treatment then--> Ceftriaxone, cefotaxime)
optreatmentorunimmunized=c("ceftriaxone","cefotaxime")

#apply guidelines
maxabxchoice<-maxabxchoice %>%
  mutate(correctinitialabx= if_else(servicelocation=="Inpatient"&(finalcorrectinitialabx==2 |finalcorrectinitialabx==4)&(Med %in% optreatmentorunimmunized),1,correctinitialabx,correctinitialabx))

#max per patient   
maxabxchoice<-maxabxchoice %>% group_by(CSN) %>% 
  mutate(finalcorrectinitialabx= max(correctinitialabx)) %>%
  ungroup()


#summary of correct vs incorrectn vs treatment failure choices (0=incorrect, 1=correct, 2=outpatient treatmentfailure
#summary of correct vs incorrect choices
summarytable3<-maxabxchoice %>%
  group_by(correctinitialabx)%>%
  summarise(count=n_distinct(CSN))%>%
  ungroup()

rowlabesl<-c("Likely Incorrect","Correct","Outpatient Treatment Failure","Aspiration Pneumonia")

cbind(rowlabesl,summarytable3)

aggregate(CSN ~ label, maxabxchoice, function(x) length(unique(x))) 
```

