---
title: "appropriate_abx_public"
author: "Julie Yarahuan"
date: "2022-09-08"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE}
#Getting started: 
## The only places that you will need to modify in the code are in the code chunks under headers which say "INPUT NEEDED". Switch to outline form in your RMD to easily see these
## The only lines within the "INPUT NEEDED" Sections that need to be edited by you will be called out with **input needed here** immediately above the variable
```


# Define Variables

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. 

```{r,echo=FALSE, message=FALSE}
#Load Libraries
library(tidyr)
library(tibble)
library(dplyr)
library(readr)
library(lubridate)
library(stringr)
library(data.table)
library(tinytex)

```

##Define Variables: INPUT NEEDED
```{r,echo=FALSE, message=FALSE}
#File path variable: Input your file path here. See below for column order info. 
#**input needed here**
abxfile="FilePathHere"

#column names for antibiotics table, make sure your data is in the below order. The column NAMES do not matter. All columns are renamed, so only the column ORDER is important. 
## 1=contact serial number (encounter ID)
## 2=MRN
## 3=date of birth
## 4= Admit DateTime
## 5=Discharge Service Abbreviation
## 6= Discharge Service description # the service the patient was discharged from
## 7=Service on a given date (service patient on on a given date (below))
## 8= date of service (date that patient on a given service)
## 9=ICD10 code
## 10=ICD 10 description
## 11=Medication Order DateTime
## 12=Medication
## 13=allergy substance 
## 14=allergy type (e.g food vs drug allergy)



#file path for output of exclusions to review. Rename each of the following paths and give the file a name that would be meaninful to you. Do not change file type (e.g csv)
#**input needed here**
exclusionfile<-"Exclusions3_9.csv"
notedatacontextfile<-"NoteDataContext3_16.csv"
fullabxreviewfile<-"FullReview3_9.csv"
csnlistfile<-"CSN3_16.csv"


##Note: for help with file paths in R you can refer to this site: https://excelquick.com/r-programming/importing-data-absolute-and-relative-file-paths-in-r/
##Note: for a very easy workaround to get the file path you can click on "import dataset" in upper right of screen--> from Text (readr)-->Browse & select your file-->copy and past only the file path string into variable above
```


```{r, echo=FALSE, message=FALSE}
#Load Datasets
abxdata <- read_csv(abxfile)

#rename columns in abx data
abxdata<-abxdata %>%
  rename(CSN=1) %>%
  rename(MRN=2) %>%
  rename(DOB=3) %>%
  rename(AdmitDT=4) %>%
  rename(DischargeService=5) %>%
  rename(DischargeServiceDesc=6) %>%
  rename(BeddedService=7)%>%
  rename(BeddedDate=8)%>%
  rename(ICD=9) %>%
  rename(ICDdesc=10) %>%
  rename(MedDT=11) %>%
  rename(Med=12) %>%
  rename(Allergy=13) %>%
  rename(AllergyType=14)



```

## Define  guideline exclusions and guideline choices: INPUT NEEDED
```{r, echo=FALSE}
#Select services of interest
#**input needed here**
servicesofinterest<-c('PMA','PMB','PMC','PMD','CCS','PUL','EMR','ADL') 

#services to exclude: this is to exclude patients who were admitted first to another service (like the ICU or a surgical service) and then transferred to PHM
#**input needed here**
prePHMservicetoexclude<-c("Intermediate Care Program","Medicine Critical Care","Medical Intensive Care")

#Minimum age allowed (e.g. this guideline excludes children less than 3 months old)
#**input needed here**
minagemonths=3

#Co-morbid conditions to exclude (make all lowercase). This will use the below phrases to search ICD 10 descriptions. It will match if the description contains the phrase. For example: "lupus" will match the ICD10 code for lupus nephritis, and systemic lupus erythematosus. 
#**input needed here**
exclusiondiagnoses=c("cystic fibrosis","other specified congenital malformations of respiratory system","sickle-cell","hb-ss","sickle cell","leukemia","lymphoma",'neoplasm','transplant',"tuberculosis",
                     "tracheostomy","end stage renal","immunodeficiency",'immunodeficiencies','hypogammaglobulinemia','di george',
                     "rheumatoid arthritis","lupus","muscular dystrophy","spinal muscular atrophy","hypoplastic left heart syndrome",
                     "pulmonary hypertension","cardiomyopathy","heart failure")


#filter out some common infections based on ICD prefixes
#AO: exclude infectious colitis, "B0*":exclude many viral infections, "J95.851":exclude ventilator assc PNA, "M00*":septic arthritis,  "M01*": septic arthritis,"N39*": exclude UTI
#**input needed here**
exclusioninfections<-c(glob2rx("A0*"),glob2rx("B0*"),"J95.851",glob2rx("M00*"),glob2rx("M01*"),glob2rx("N39*"))

#similar to comorbid conditions, the below will use the following phrases to search ICD10 descriptions. 
exclusioninfections2=c("abscess","cellulitis", "osteomyelitis",
                     "mastoiditis","lymphangitis","sinusitis","streptococcal pharyngitis","impetigo","lymphadenitis",
                     "omphalitis","aspiration","arthritis due to other bacteria","myositis","sinusitis",'pyelonephritis')

#filter out certain antimicrobials
#**input needed here**
exclusionantimicrobials<- c('oseltamivir','bacitracin','nystatin','valacyclovir','hydroxychloroquine','erythromycin')

#identify list of pneumonia ICD codes of interest (these are the codes of pneumonia that you are INCLUDING in your table)
#**input needed here**
pnalist<-c('J13', 'J13', 'J14', 'J14', 'J15.0', 'J15.0', 'J15.1', 'J15.1', 'J15.20', 'J15.20', 'J15.211', 'J15.211', 'J15.212', 'J15.212', 'J15.29', 'J15.29', 'J15.3', 'J15.3', 'J15.4', 'J15.4', 'J15.5', 'J15.5', 'J15.6', 'J15.6', 'J15.7', 'J15.7', 'J15.8', 'J15.8', 'J15.9', 'J15.9', 'J16.0', 'J16.0', 'J16.8', 'J17', 'J18.0', 'J18.1', 'J18.8', 'J18.9', 'J85.1', 'J95.851', 'J95.851')


#define first line choices 
#**input needed here**
firstline=c("amoxicillin","ampicillin") #first line is ampicillin or amoxicillin
#**input needed here**
PCNallergy=c("amoxicillin","ampicillin",'penicillin','augmentin','penicillins','amoxicillin-clavulanate') #list of meds that count as PCN allergy
#**input needed here**
ifPCNallergy=c("ceftriaxone","clindamycin","cefdinir",'levofloxacin') #if PCN allergy then, ceftriaxone, clindamycin, cefdinir, levofloxacin
#**input needed here**
cephalosporinallergy=c("cefdinir","cefotaxime","ceftriaxone","cephalexin","cephalosporins","cefuroxime","cefazolin","cefadroxil","cefotetan","cefoxitin","ceftazidime","cefpodoxime","cefixime","cefepime","ceftaroline") #list of meds that count as cephalosporin allergy
#**input needed here**
ifcephalosporinallergy=c("levofloxacin","clindamycin") #if cephalosporin allergy then, levofloxacin, clindamycin
#**input needed here**
atypicalfirstline="azithromycin" #first line atypical coverage is azithromycin
#**input needed here**
azithroallergy="azithromycin"
#**input needed here**
atypicalwithazithroallergy=c("levofloxacin","doxycycline") #if azithromycin allergy then levofloxacin or doxycycline


#if over 4 then azithromycin is potentially appropriate
#**input needed here**
ageforazithro=4 

#define guidelinechoices (if unimmunized then--> Ceftriaxone, cefotaxime, cefdinir, levofloxacin)
#**input needed here**
treatmentunimmunized=c("ceftriaxone","cefotaxime","cefdinir","levofloxacin")

```

```{r, echo=FALSE}
################################# Clean Dataset ########################################

#reformat dates for admit DT and DOB
abxdata$AdmitDT<-mdy_hms(abxdata$AdmitDT, truncated = 1)
abxdata$DOB<-mdy(abxdata$DOB)
abxdata$MedDT<-mdy_hms(abxdata$MedDT)

#Calculate age in years at time of admission
abxdata <- abxdata %>%
  mutate(age=year(AdmitDT)-year(DOB))

#calculate age in months at time of admission
abxdata<-abxdata %>%
  mutate(agemonths=interval(DOB,AdmitDT) %/% months(1))

#make text columns lowercase
abxdata$ICDdesc=tolower(abxdata$ICDdesc)
abxdata$Med=tolower(abxdata$Med)
abxdata$Allergy=tolower(abxdata$Allergy)

#if allergy is na then fill in with "no known medication allergies"
abxdata<-abxdata%>%
  mutate(Allergy=if_else(is.na(Allergy),"no known medication allergies",Allergy))
```


##Count starting encounters
```{r, echo=FALSE}
#Count unique starting CSNs
csnunique <- unique(abxdata$CSN)
paste("Total starting encounters: ",length(csnunique))

```

# Output

## Summary of reasons for exlusion
```{r, echo=FALSE}
################################# Apply guideline exclusions ########################################
#eliminate patients less than your min exclusion age
abxdata <- abxdata %>%
  mutate(reasonforexclusions= if_else(
    agemonths>=minagemonths,"Included","Below min age"))


#select services of interest (discharge service=PHM)
abxdata <- abxdata %>%
  mutate(reasonforexclusions= if_else(
    (DischargeService %in% servicesofinterest) &reasonforexclusions=="Included","Included",
    if_else(reasonforexclusions!="Included",reasonforexclusions,"Excluded service")))

#filter out patients admitted to ICU/ICP before PHM
excludeservicelist<-abxdata %>%
  select(CSN) %>%
  filter(abxdata$BeddedService %like% paste(prePHMservicetoexclude,collapse = "|"))

#apply exclude list
abxdata <- abxdata %>%
  mutate(reasonforexclusions= if_else(
    CSN %in% excludeservicelist$CSN & reasonforexclusions=="Included","Excluded service",
    if_else(reasonforexclusions!="included",reasonforexclusions,"Included")))

#filter diagnoses that have any of the excluded diagnoses
excludelist<-abxdata %>%
  select(CSN) %>%
  filter(str_detect(abxdata$ICDdesc, paste(exclusiondiagnoses,collapse='|')))

abxdata <- abxdata %>%
  mutate(reasonforexclusions= if_else(
    CSN %in% excludelist$CSN & reasonforexclusions=="Included","Excluded diagnosis",
    if_else(reasonforexclusions!="Included",reasonforexclusions,"Included")))

#filter diagnoses that have any other common infections
excludeinfectionlist<-abxdata %>%
  select(CSN) %>%
  filter(abxdata$ICD %like% paste(exclusioninfections,collapse = "|"))

#apply exclude list
abxdata <- abxdata %>%
  mutate(reasonforexclusions= if_else(
    CSN %in% excludeinfectionlist$CSN & reasonforexclusions=="Included","Excluded other common infection",
    if_else(reasonforexclusions!="included",reasonforexclusions,"Included")))

#common infections exclusions list 2
excludeinfectionlist2<-abxdata %>%
  select(CSN) %>%
  filter(str_detect(abxdata$ICDdesc, paste(exclusioninfections2,collapse='|')))

abxdata <- abxdata %>%
  mutate(reasonforexclusions= if_else(
    CSN %in% excludeinfectionlist2$CSN & reasonforexclusions=="Included","Excluded other common infection",
  if_else(reasonforexclusions!="Included",reasonforexclusions,"Included")))


#exclude selected antimicrobials from med list (first we ID patients who received any antimicrobial that is NOT on the excluded list)
noexcludemedlist<-abxdata%>%
  select(CSN)%>%
  filter(!abxdata$Med %like% paste(exclusionantimicrobials,collapse = "|"))  
noexcludemedlist<-unique(noexcludemedlist)

#get a list of all unique CSNs to compare to
allcsn<-abxdata %>%
  select(CSN)
allcsn<-unique(allcsn)

#Next we find only patients who received ONLY an antimicrobial that is on the excluded list
excludemedlist<-setdiff(allcsn,noexcludemedlist)

#label "did not receive abx" only for patients who ONLY received one of the excluded medications and no other antimicrobial
abxdata <- abxdata %>%
  mutate(reasonforexclusions= if_else(
    CSN %in% excludemedlist$CSN & reasonforexclusions=="Included","Did not receive antibiotic",
  if_else(reasonforexclusions!="Included",reasonforexclusions,"Included")))

#filter out meds that are in the excluded list for patients who recieved other antimicrobials
abxdata<-abxdata%>%
  filter((!Med %like% paste(exclusionantimicrobials,collapse="|")|(CSN %in% excludemedlist$CSN)))

#pna subset
pnasub<-abxdata %>%
  select(CSN) %>%
  filter(abxdata$ICD %in% pnalist)

abxdata <- abxdata %>% 
  mutate(reasonforexclusions= if_else(CSN %in% pnasub$CSN & reasonforexclusions=="Included","Included",
                                      if_else(reasonforexclusions!="Included",reasonforexclusions,"No common pnuemonia ICD code"))) 


exclusionsummarytable<-setNames(aggregate(CSN ~ reasonforexclusions, abxdata, function(x) length(unique(x))),c("Reason For Exclusion","Count Excluded Encounters"))
exclusionsummarytable<-exclusionsummarytable %>%
  arrange(desc(`Count Excluded Encounters`))

grandtotal<-c("Grand Total",length(unique(abxdata$CSN)))
exclusionsummarytable<-rbind(exclusionsummarytable,grandtotal)


exclusionsummarytable


```

```{r, echo=FALSE,message=FALSE, results=FALSE}
################################# Apply guidelines to assess antibiotic choice ########################################
#find first abx ordered at encounter
abxchoice <- abxdata %>%
  group_by(CSN) %>%
  filter(
    MedDT== min(MedDT, na.rm = T) & reasonforexclusions=="Included") 

#create a blank column to store reasons abx are classified as incorrrect: 
abxchoice <- abxchoice %>%
  add_column(label = NA)%>%
  add_column(labeldetailed=NA)

#correct based on defined guidelines in definitions section
abxchoice<-abxchoice %>%
  ungroup()%>%
  mutate(correctinitialabx=case_when(
    Med %in% firstline ~1,
    (Allergy %in% PCNallergy & Med %in% ifPCNallergy)~1,#if allergy to pencillin, then Ceftriaxone, clinda, cefdinir or levofloxacin are correct
    (Allergy %in% cephalosporinallergy & Med %in% ifcephalosporinallergy)~1,#if cephalosporin allergy then clinda or levofloxacin
    (age>ageforazithro & Med==atypicalfirstline)~1,#if over 4 then azithromycin correct,
    (age>ageforazithro & Allergy==azithroallergy & Med %in% atypicalwithazithroallergy)~1, #if over 4 and has azithro allergy
    TRUE~0
  ))

#add reasons for being correct
abxchoice<-abxchoice %>%
  ungroup()%>%
    mutate(label=case_when(
    correctinitialabx==1 ~"Correct",TRUE
    ~"Needs review"
  ))%>%
  mutate(labeldetailed=case_when(
    Med %in% firstline ~"First line choice",
    (Allergy %in% PCNallergy & Med %in% ifPCNallergy)~"Correct substitute i/s/o PCN allergy",
    (Allergy %in% cephalosporinallergy & Med %in% ifcephalosporinallergy)~"Correct substitute i/s/o cephalosporin allergy",
    (age>ageforazithro & Med==atypicalfirstline)~"Appropriate age for Azithro use",
    (age>ageforazithro & Allergy==azithroallergy & Med %in% atypicalwithazithroallergy)~"Appropriate age for azithro with appropriate azithro subs i/s/o   Azithro allergy",
    TRUE~"Needs review"
  ))

#take max of choice per patietn (this eliminates rows that look false [e.g. patient with multiple drug allergies, recorded as 'incorrect' for one allergy but correct for another])
maxabxchoice<-abxchoice %>% group_by(CSN) %>% top_n(1,correctinitialabx)
  
# Define the desired order of levels
level_order <- c("First line choice", 
                  "Correct substitute i/s/o PCN allergy",
                  "Correct substitute i/s/o cephalosporin allergy",
                  "Appropriate age for Azithro use",
                  "Appropriate age for azithro with appropriate azithro subs i/s/o   Azithro allergy",
                  "Needs review")

# Convert labeldetailed to a factor with the desired order of levels
maxabxchoice$labeldetailed <- factor(maxabxchoice$labeldetailed, levels = level_order)


#now take only the first "level" that appears in the datafram. That way if patient would otherwise match on two labels (correct i/s/o PCN allergy & correct i/s/o cephalosporin allergy), now we only have one unique match per patient

# keep only the minimum value of the label_detailed column for each CSN
maxabxchoice <- maxabxchoice %>%
  group_by(CSN) %>%
  slice_min(order = labeldetailed)






```



## Summary of Correct/Incorrect Orders with Labels [Pre-NLP Analysis]
```{r,echo=FALSE, message=FALSE}
###################### Display Part 1 results ##################

#summary of correct vs incorrect choices
summarytable1<-maxabxchoice %>%
  group_by(label)%>%
  summarise(count=n_distinct(CSN))%>%
  ungroup()

grandtotal<-c("Grand Total",length(unique(maxabxchoice$CSN)))
summarytable1<-rbind(summarytable1,grandtotal)

summarytable1

#Summary of labeldetailed
summarytable2<-maxabxchoice%>%
  group_by(label,labeldetailed)%>%
  summarise(count=n_distinct(CSN))%>%
  ungroup()%>%
  arrange(label,desc(count))

grandtotal2<-c("Grand Total","",length(unique(maxabxchoice$CSN)))
summarytable2<-rbind(summarytable2,grandtotal2)
summarytable2
```

#Start NLP

##Export List of CSN's to get Note Text:Outside R Data Collection Step
```{r, echo=FALSE}
################################# get list of all remaining patients who are incorrect for NLP below ########################################

#get CSN list of remaining patients who need review. You will export this list of CSNs, get all the note text (see next code chunk) to then run the NLP
csnlist<-maxabxchoice[maxabxchoice$correctinitialabx == 0, ]
csnlist<-unique(csnlist$CSN)

#save CSN list as CSV
write.csv(csnlist,csnlistfile, row.names = FALSE)
```

##Import Note Data: INPUT NEEDED
```{r, echo=FALSE}
#File path variable: Input your file path to note contents file here
#**input needed here**
notefile="/Volumes/CI Program Faculty/+Fellow Franchise/Yarahuan/Protected/Notes with Note Contents to delete.csv"
#NoteData Columns
#1=CSN, 2=NoteDT, 3=NoteType, 4=NoteText
```

##Create NLP Phrases: INPUT NEEDED
```{r, echo=FALSE,message=FALSE}
################################# Create your NLP phrases ########################################################
#NLP notes: 
## General structure of each regex: 1. (negative look behind asssertion-- things that should NOT PRECEED our phrases) 2. (matching phrases) 3.(negative look ahead assertion-- things that should NOT FOLLOW our phrases)
## \\s+ is used to match any whitespace characters (such as spaces, tabs, or newlines) that might appear between words in the phrases being matched.
## (?i) - This is a flag that makes the search case-insensitive. It means that the regex will match both uppercase and lowercase letters.
## (?:) non-capturing group to group together common parts of the phrases. 
## (?<!(unlikely|less likely|did not) - This is a negative lookbehind assertion, which means that the regex will only match if the previous word is not one of the listed words followed by a space. This part of the regex is checking for words that indicate that treatment failure is unlikely or did not occur.
## Similar to above, you can also have a negative lookahead assertion, which means the phrases cannot be followed by any of the phrases in that assertion. You will find that assertion after the main phrases
## (?:) is used to create a non-capturing group within a regular expression. This means that the group is used for matching purposes, but the content that it matches is not captured and stored as a separate group for later use in the regular expression or in the code that uses the regular expression. For example:  history of(?: an)?, the (?:) syntax is used again to create a non-capturing group. The group contains the string " an" with a ? after it, which means that the an is optional. This group is preceded by the string "history of".


#treatment failure phrases
#**input needed here**
treatmentfailurephrases<-"(?i)(?<!(unlikely|less likely|did not|no|has no|didn't|hasn't|has not yet|hasn't yet|did not yet|didn't yet)\\s)(treatment failure|failed outpatient\\s+(?:treatment|therapy|antibiotics|antibiotic|amoxicillin|HD amoxicillin|high-dose amoxicillin|high dose amoxicillin|HD amoxicillin)|failure of outpatient\\s+(?:treatment|therapy|antibiotics|antibiotic|amoxicillin|HD amoxicillin|high-dose amoxicillin|high dose amoxicillin)|despite\\s+(?:treatment|therapy|antibiotics|antibiotic|amoxicillin|HD amoxicillin|high-dose amoxicillin|high dose amoxicillin |outpatient\\s+(?:treatment|therapy|antibiotics|antibiotic|amoxicillin|HD amoxicillin|high-dose amoxicillin|high dose amoxicillin))|(?:in spite of |s p|sp|s/p)\\s+outpatient\\s+(?:treatment|therapy|antibiotics|antibiotic|amoxicillin|HD amoxicillin|high-dose amoxicillin|high dose amoxicillin)|started on\\s+(?:treatment|therapy|antibiotics|antibiotic|amoxicillin|HD amoxicillin|high-dose amoxicillin|high dose amoxicillin)\\s+as\\s+(?:outpatient|an outpatient))"



#aspiration phrases
#**input needed here**
aspirationphrases="(?i)(?<!unlikely|less likely|is unlikely|is less likely|no|not|not likely|reassuring against(?: an)?|no concern for(?: an)?|am not concerned for(?: an)?|no history (?: of)?(?: an)?|no recent episodes of(?: an)?|history of(?: an)?|no clear|no recent|no obvious|recurrent|denies(?: an)?|low suspicion for(?: an)?|did not witness(?: an)?|multiple)\\s+(aspiration pneumonia|aspiration PNA|aspiration event|presumed aspiration|empirically for aspiration|cover for aspiration|treat for aspiration|aspiration coverage|aspiration treatment)(?!\\s+unlikely|\\s+less likely|\\s+is unlikely|\\s+is less likely|\\s+not likely|\\s+is not likely|\\s+is not needed|\\s+is not necessary)"

#immunization phrases
#**input needed here**
unimmunizedphrases="(?i)unimmunized|unvaccinated|partially\\s+(vaccinated|immunized)|not\\s+(immunized|fully immunized|up to date on immunizations|vaccinated|fully vaccinated)|immunizations\\s+(not up to date|are not up to date|incomplete)|incomplete\\s+(immunizations|vaccinations)"

#critical illness phrases
#**input needed here**
criticalillnessphrases="(?i)(?<!no|not|not in|no signs of|no sign of|not concerned for|no concern for|reassured against|reassuring against|no evidence of|not suggestive of|risk factors for|at risk for|high risk for|no concern for)\\s+(shock|critically ill|sepsis|critical illness|septic)(?!\\s(screening|screen))"

#pleural effusion
#**input needed here**
pleuraleffusionphrases="(?i)(?<!no|no signs of(?: a)?|no sign of(?: a)?|not concerned for(?: a)?|no concern for(?: a)?|reassured against(?: a)?|reassuring against(?: a)?|no evidence of(?: a)?|not suggestive of(?: a)?|small to)\\s+((moderate|large|complex|loculated)\\s+(?:right\\s|right sided\\s|right-sided\\s|left\\s|left sided\\s|left-sided\\s)?(pleural|parapneumonic)\\s+effusion)(?!\\s+was not visualized)"

#immunocomrpomised
#**input needed here**
immunesuppressionphrases="(?i)(?<!no|not|no signs of(?: an)|no sign of(?: an)|not concerned for(?: an)|no concern for(?: an)|reassured against(?: an)|reassuring against(?: an)|no evidence of(?: an)|not suggestive of(?: an)|no (history|hx) of(?: an)|never (?: been)? evaluated for(?: an)|no family (history|hx) of(?: an)?|no fhx of(?: an)?|negative for(?: an))\\s+(immunosuppressed|immunocompromised|immune compromised|immune suppressed|immunodeficiency|immunodeficient|immune deficiency|immune deficient)"

```



```{r,echo=FALSE, message=FALSE}
################################# Start NLP ########################################################

#Load Datasets
notedata <- read_csv(notefile)

#rename columns
notedata<-notedata %>%
  rename(CSN=1) %>%
  rename(NoteDT=2) %>%
  rename(NoteType=3) %>%
  rename(NoteText=4)

#make text columns lowercase
notedata$NoteType=tolower(notedata$NoteType)

```

```{r, echo=FALSE}
#count unique patients to make sure matches from those needing review in part 1
csnunique <- unique(notedata$CSN)
paste("Total starting encounters: ",length(csnunique))
```


## Apply NLP phrases
```{r,echo=FALSE,message=FALSE}

################################# Apply NLP phrases ########################################################

#apply treatment failure phrases
treatmentfailure<-notedata %>%
  select(CSN) %>%
  filter(str_detect(notedata$NoteText, treatmentfailurephrases))

#apply aspiration phrases
aspiration<-notedata %>%
  select(CSN) %>%
  filter(str_detect(notedata$NoteText, aspirationphrases))

#apply unimmunized phrases
unimmunized<-notedata %>%
  select(CSN) %>%
  filter(str_detect(notedata$NoteText, unimmunizedphrases))

#apply critical illness phrases
criticalillness<-notedata %>%
  select(CSN) %>%
  filter(str_detect(notedata$NoteText, criticalillnessphrases))

#apply pleural effusion phrases
pleuraleffusion<-notedata %>%
  select(CSN) %>%
  filter(str_detect(notedata$NoteText, pleuraleffusionphrases))

#apply immune suppressed phrases
immunesuppressed<-notedata %>%
  select(CSN) %>%
  filter(str_detect(notedata$NoteText, immunesuppressionphrases))


#get all CSNs from above lists
treatmentfailurecsnlist<-unique(treatmentfailure$CSN)
aspirationcsnlist<-unique(aspiration$CSN)
unimmunizedcsnlist<-unique(unimmunized$CSN)
criticalillnesscsnlist<-unique(criticalillness$CSN)
pleuraleffusioncsnlist<-unique(pleuraleffusion$CSN)
immunosuppressedcsnlist<-unique(immunesuppressed$CSN)

#convert labeldetailed back to character vector
maxabxchoice$labeldetailed<-as.character(maxabxchoice$labeldetailed)

#assign labels
maxabxchoice <- maxabxchoice %>%
  ungroup() %>%
  mutate(correctinitialabx = if_else(
    maxabxchoice$CSN %in% treatmentfailurecsnlist | maxabxchoice$CSN %in% aspirationcsnlist | 
      maxabxchoice$CSN %in% unimmunizedcsnlist | maxabxchoice$CSN %in% pleuraleffusioncsnlist | 
      maxabxchoice$CSN %in% immunosuppressedcsnlist | maxabxchoice$CSN %in% criticalillnesscsnlist,
    2,
    correctinitialabx
  )) %>%
  mutate(label = if_else(
    maxabxchoice$CSN %in% treatmentfailurecsnlist,
    "Needs review",
    if_else(
      maxabxchoice$CSN %in% aspirationcsnlist,
      "Needs review",
      if_else(
        maxabxchoice$CSN %in% unimmunizedcsnlist,
        "Needs review",
        if_else(
          maxabxchoice$CSN %in% pleuraleffusioncsnlist,
          "Needs review",
          if_else(
            maxabxchoice$CSN %in% immunosuppressedcsnlist,
            "Needs review",
            if_else(
              maxabxchoice$CSN %in% criticalillnesscsnlist,
              "Needs review",
              label
            )
          )
        )
      )
    )
  )) %>%
  mutate(labeldetailed = if_else(
    maxabxchoice$CSN %in% treatmentfailurecsnlist,
    "Outpatient treatment failure",
    if_else(
      maxabxchoice$CSN %in% aspirationcsnlist,
      "Aspiration pneumonia",
      if_else(
        maxabxchoice$CSN %in% unimmunizedcsnlist,
        "Unimmunized",
        if_else(
          maxabxchoice$CSN %in% pleuraleffusioncsnlist,
          "Large/Complicated Pleural Effusion",
          if_else(
            maxabxchoice$CSN %in% immunosuppressedcsnlist,
            "Immunosuppressed",
            if_else(
              maxabxchoice$CSN %in% criticalillnesscsnlist,
              "Critically ill",
              labeldetailed
            )
          )
        )
      )
    )
  ))




#apply guidelines for unimmunized patient
maxabxchoice<-maxabxchoice %>%
  mutate(correctinitialabx= if_else((labeldetailed=="Unimmunized")&(Med %in% treatmentunimmunized),1,correctinitialabx,correctinitialabx))%>%
  mutate(label= if_else((labeldetailed=="Unimmunized")&(Med %in% treatmentunimmunized),"Correct",label,label))%>%
  mutate(labeldetailed= if_else((labeldetailed=="Unimmunized")&(Med %in% treatmentunimmunized),"Correct substitute for unimmunzed",labeldetailed,labeldetailed))



```

##Summary with NLP 
```{r, echo=FALSE}
#summary post-NLP 
#max per patient   
maxabxchoice<-maxabxchoice %>% group_by(CSN) %>% 
  mutate(correctinitialabx= max(correctinitialabx)) %>%
  ungroup()

maxabxchoice %>%
  group_by(label,labeldetailed)%>%
  summarise(count=n_distinct(CSN))%>%
  ungroup()%>%
  arrange(label,desc(count))
```
#Generate output for review
```{r, echo=FALSE}
############ get phrase's ID and note context around phrases ########

######combine all phrases and CSN's#######3
allNLP_phrases<-c(paste(treatmentfailurephrases,collapse = "|"),paste(aspirationphrases,collapse = "|"),paste(unimmunizedphrases,collapse = "|"),paste(criticalillnessphrases,collapse = "|"),paste(pleuraleffusionphrases,collapse = "|"),paste(immunesuppressionphrases,collapse = "|"))
all_CSN_NLP_list<-c(treatmentfailurecsnlist,aspirationcsnlist,unimmunizedcsnlist,criticalillnesscsnlist,pleuraleffusioncsnlist,immunosuppressedcsnlist)
#make new columns
notedata<-notedata%>%
  mutate(startposition="NA") %>%
  mutate(endposition="NA")%>%
  mutate(context="NA")

#show phrase captured in column
notedata<-notedata %>%
  ungroup()%>%
  mutate(identifiedphrase= if_else(notedata$CSN %in% all_CSN_NLP_list
                                  ,str_extract(notedata$NoteText,paste(allNLP_phrases,collapse = "|")),""))

####for loop 
for(i in 1:nrow(notedata)) {       # for-loop over rows
 notedata$startposition[i]<-((str_locate(notedata$NoteText[i],notedata$identifiedphrase[i])[,1])-120)
 notedata$endposition[i]<-((str_locate(notedata$NoteText[i],notedata$identifiedphrase[i])[,2])+120)
 notedata$context[i]=str_sub(notedata$NoteText[i],start=if_else((notedata$startposition[i])>0,(notedata$startposition[i]),"0"),end=(notedata$endposition[i]))
}
#labeldetailed added to notedata
labeldetailedlist <- maxabxchoice %>%
   select(CSN,labeldetailed, Med)

#join note data with maxabx data to get labels
notedata <- merge(x=notedata,y=labeldetailedlist, 
             by="CSN")


```

##Write output to a file
```{r, echo=FALSE}

#######Show summary statistics for phrases
notedatasummary<-notedata %>%
  group_by(identifiedphrase)%>%
  summarise(count=n_distinct(CSN))%>%
  ungroup()%>%
  arrange(desc(count))

####Creat table for easy phrase/context manual review
notedatacontext<-notedata%>%
  select(CSN,Med,NoteDT,NoteType,identifiedphrase,context,labeldetailed)

#delete redundant rows (example blank rows, when other rows have phrase ID & context)
notedatacontext <- notedatacontext %>%
  filter(!(is.na(identifiedphrase) | identifiedphrase == "") | labeldetailed == "Needs review") %>%
  distinct(CSN, .keep_all = TRUE)


#export notes with note content for review
write.csv(notedatacontext,notedatacontextfile,row.names = FALSE)

```

##Additional Optional Output for Review
```{r, echo=FALSE}
#Additional output for review

#review exclusions
write.csv(abxdata,exclusionfile, row.names = FALSE)

#join main table with NLP note context-- this will pull all patient data + note data with context
fullreview= maxabxchoice %>% left_join(notedatacontext,by="CSN")
write.csv(fullreview,fullabxreviewfile, row.names = FALSE)
```
